---
title: "Wave Patterns Before, During and After Hurricanes"
author: "Alexandria Cerpovicz"
subtitle: Wave height and direction before and after Hurricanes Irma and Maria in
  St. John, USVI
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---

# Introduction

## Problem / Question
What was the difference in wave heights before during and after hurricanes Irma and Maria passed St. John, USVI in 2017? What was the difference between the wave height between the two hurricanes. 

#Materials and Methods

## Proposed Methods

This study will use the following packages to answer the questions above. 

1. To import the data from the buoy:
  *rnoaa, to import buoy data
  *ncdf4, to easily read in the buoy dataset
  *leaflet, used to create interactive map of buoy locations

2. To import hurricane data:
  *sf, to encode spatial vector data in a dataframe
  *tidyverse
  *rnoaa
  *spData
  *data(world)
  *data(us_states)

3. Wave rose digrams: to show the differences in wave heights before durning and after the two hurricanes:
  * ggplot2, to make wave rose diagrams
  * RColorBrewer, to make wave rose diagrams
  
4. Compare wave heights between two hurricanes:
  * plotly, to create a time series line plot


# Import Data Sets
```{r, message=F, warning=F}
library(rnoaa)
library(ncdf4)
library(leaflet)
knitr::opts_chunk$set(cache=TRUE)
```

## NOAA Buoy Stations Information and Stations Map
```{r}
allbuoystations <- buoy_stations()
leaflet(data = na.omit(allbuoystations)) %>%
  leaflet::addTiles() %>%
  leaflet::addCircles(~lon, ~lat, opacity = 0.5)
```

## NOAA St. John Station Map
```{r}
stjohnbuoy <- filter(allbuoystations, station==41052)
head(stjohnbuoy)
leaflet(data = na.omit(stjohnbuoy)) %>%
  leaflet::addTiles() %>%
  leaflet::addCircles(~lon, ~lat, opacity = 0.5)

```


## Standard meteorological data at buoy 41052
```{r, echo= FALSE, results= 'hide', message=F, warning=F}
st_john_data <- buoy(dataset = 'stdmet', buoyid = 41052, year = 2017)
head(st_john_data)
```

## Hurricane Data Import

```{r}
library(sf)
library(tidyverse)
library(ggmap)
library(rnoaa)
library(spData)
data(world)
data(us_states)
```

```{r}
hurricanes <- storm_shp(basin = "NA") %>%
storm_shp_read() %>%
st_as_sf()
hurricanes2 <- filter(hurricanes, year == 2017) %>%
mutate_if(is.numeric, function(x) ifelse(x==-999.0,NA,x)) 
table(hurricanes$year)
head(hurricanes2)
```

# Results

## Severe Weather in St. John During Hurricane Season 1995

swdi(dataset='warn', startdate = '19950601', enddate = '19951101', station = '41052')

## Wave Height Before During and After Hurricane

### Wave rose diagram of week before hurricanes

```{r}
head(st_john_data)
install.packages("ggplot2")
library(ggplot2)
install.packages("RColorBrewer")
library(RColorBrewer)

plot.windrose <- function(data,
                      spd,
                      dir,
                      spdres = 2,
                      dirres = 30,
                      spdmin = 2,
                      spdmax = 20,
                      spdseq = NULL,
                      palette = "YlGnBu",
                      countmax = NA,
                      debug = 0){


# Look to see what data was passed in to the function
  if (is.numeric(spd) & is.numeric(dir)){
    # assume that we've been given vectors of the speed and direction vectors
    data <- data.frame(spd = spd,
                       dir = dir)
    spd = "spd"
    dir = "dir"
  } else if (exists("data")){
    # Assume that we've been given a data frame, and the name of the speed 
    # and direction columns. This is the format we want for later use.    
  }  

  # Tidy up input data ----
  n.in <- NROW(data)
  dnu <- (is.na(data[[spd]]) | is.na(data[[dir]]))
  data[[spd]][dnu] <- NA
  data[[dir]][dnu] <- NA

  # figure out the wind speed bins ----
  if (missing(spdseq)){
    spdseq <- seq(spdmin,spdmax,spdres)
  } else {
    if (debug >0){
      cat("Using custom speed bins \n")
    }
  }
  # get some information about the number of bins, etc.
  n.spd.seq <- length(spdseq)
  n.colors.in.range <- n.spd.seq - 1

  # create the color map
  spd.colors <- colorRampPalette(brewer.pal(min(max(3,
                                                    n.colors.in.range),
                                                min(9,
                                                    n.colors.in.range)),                                               
                                            palette))(n.colors.in.range)

  if (max(data[[spd]],na.rm = TRUE) > spdmax){    
    spd.breaks <- c(spdseq,
                    max(data[[spd]],na.rm = TRUE))
    spd.labels <- c(paste(c(spdseq[1:n.spd.seq-1]),
                          '-',
                          c(spdseq[2:n.spd.seq])),
                    paste(spdmax,
                          "-",
                          max(data[[spd]],na.rm = TRUE)))
    spd.colors <- c(spd.colors, "grey50")
  } else{
    spd.breaks <- spdseq
    spd.labels <- paste(c(spdseq[1:n.spd.seq-1]),
                        '-',
                        c(spdseq[2:n.spd.seq]))    
  }
  data$spd.binned <- cut(x = data[[spd]],
                         breaks = spd.breaks,
                         labels = spd.labels,
                         ordered_result = TRUE)
  # clean up the data
  data. <- na.omit(data)

  # figure out the wind direction bins
  dir.breaks <- c(-dirres/2,
                  seq(dirres/2, 360-dirres/2, by = dirres),
                  360+dirres/2)  
  dir.labels <- c(paste(360-dirres/2,"-",dirres/2),
                  paste(seq(dirres/2, 360-3*dirres/2, by = dirres),
                        "-",
                        seq(3*dirres/2, 360-dirres/2, by = dirres)),
                  paste(360-dirres/2,"-",dirres/2))
  # assign each wind direction to a bin
  dir.binned <- cut(data[[dir]],
                    breaks = dir.breaks,
                    ordered_result = TRUE)
  levels(dir.binned) <- dir.labels
  data$dir.binned <- dir.binned

  # Run debug if required ----
  if (debug>0){    
    cat(dir.breaks,"\n")
    cat(dir.labels,"\n")
    cat(levels(dir.binned),"\n")       
  }  

  # deal with change in ordering introduced somewhere around version 2.2
  if(packageVersion("ggplot2") > "2.2"){    
    cat("Hadley broke my code\n")
    data$spd.binned = with(data, factor(spd.binned, levels = rev(levels(spd.binned))))
    spd.colors = rev(spd.colors)
  }

  # create the plot ----
  p.windrose <- ggplot(data = data,
                       aes(x = dir.binned,
                           fill = spd.binned)) +
    geom_bar() + 
    scale_x_discrete(drop = FALSE,
                     labels = waiver()) +
    coord_polar(start = -((dirres/2)/360) * 2*pi) +
    scale_fill_manual(name = "Wind Speed (m/s)", 
                      values = spd.colors,
                      drop = FALSE) +
    #theme_bw() +
    theme(axis.title.x = element_blank(),
          #panel.border = element_rect(colour = "blank"),
          panel.grid.major = element_line(colour="grey65"))

  # adjust axes if required
  if (!is.na(countmax)){
    p.windrose <- p.windrose +
      ylim(c(0,countmax))
  }

  # print the plot
  print(p.windrose)  

  # return the handle to the wind rose
  return(p.windrose)
}

p1 <- plot.windrose(st_john_data, spd = st_john_data$wave_height,
                   dir = st_john_data$mean_wave_dir)

```

### Wave rose diagram of each hurricane's peak wave dates
### Wave rose diagram of week after each hurricane

## Time Series Line Plot of Hurricane Wave Height Differences

```{r}
st_john_df <- (st_john_data$data)
 
st_john_hurricane_season <- st_john_df %>% mutate(month = substr(time, 6, 7)) %>% mutate(day= substr(time, 9, 10)) %>% filter(month %in% c("06","07", "08", "09", "10", "11"))
  
st_john_by_day <- st_john_hurricane_season %>% mutate(time= as.Date(paste0("2017", "-", month,"-", day))) %>% group_by(time) %>% summarize(wave_average= mean(wave_height,na.rm = T))

st_john_hurricane_plot <- st_john_by_day %>% ggplot(aes(y= wave_average, x= time)) + geom_line()
st_john_hurricane_plot
```

# Conclusions

What have you learned?  Are there any broader implications?

# References
Scott Chamberlain (2017). rnoaa: ‘NOAA’ Weather Data from R. R package version 0.7.0. https://CRAN.R-project.org/package=rnoaa
https://ropensci.org/tutorials/rnoaa_tutorial/
https://rmarkdown.rstudio.com/
https://github.com/dkahle/ggmap


